---
title: "All plots- manifold project"
author: "Borenstein Lab"
date: "March 2023"
output:
  html_document:
    css: custom_notebook_formatting.css
    toc: true
    toc_depth: 3
    df_print: paged
    code_folding: hide
---

```{r setup, include = FALSE}
rm(list=ls())

library(vegan)
library(ggarchery)
library(nlme)
library(data.table)
library(ggpubr)
library(ggpmisc)
library(scales)
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(writexl)
library(ggpubr)
library(RColorBrewer)
library(xlsx)
library(tibble)
library(extrafont)
library(showtext)
library(cowplot)
library(circlize)
library(gtools)
library(ComplexHeatmap)
font_add_google("Gafata")
showtext_auto()
```

## Variables

```{r variables}
top_path = '../data/'

file_name = paste0(top_path, 'ps_res_PCA_umap_02072023.xlsx')

sheet_abun = 'abundance'
sheet_meta = 'meta'
sheet_umap = 'umap'

menst_file = paste0(top_path, 'menst_statistics_res_PCA_umap_14102023.xlsx')

bactria_file_name = paste0(top_path, 'branch_melt_df_02072023.xlsx')
sheet_melt_df = 'melt_branch_df'


nn_file_name = paste0(top_path, 'nn_res_PCA_umap_02072023.xlsx')
nn_menst_file = paste0(top_path, 'nn_menst_statistics_res_PCA_umap_14102023.xlsx')


save_path = paste0(top_path, 'figures/')

date = '02072023'
```

## Colors pallettes

```{r}
community_colors = c('I-A' = '#9d0208', 'I-B' = '#dc2f02', 'II' = '#FFD60A', 'III-A' = '#e85d04','III-B' = '#faa307','IV-A' = '#22577a', 'IV-B' = '#5390D9', 'IV-C0' = '#2E214F', 'IV-C1' = '#63479B','IV-C2' = '#8962C7', 'IV-C3' = '#AB7DDA', 'IV-C4' = '#DEC9E9', 'V' = '#E2D6B8')

db_colors = c('temp_pyro' = '#F7C59F', 'temp_hiseq' = '#EFEFD0', 'cros_I' = '#89B1D0', 'cros_II' = '#1A659E', 'carter' = '#FA5C23', 'srin' = '#00365F')

coltypes_vec = c("text", "text", "text", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", 
                 "text", "text", "numeric", "text", "numeric", "numeric", "numeric", "numeric", "text",
                 "text", "numeric", "numeric", "numeric", "text", "numeric", "numeric", "numeric", "numeric", 
                 "text", "numeric", "text", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric")
```

```{r}
get_data = function(path, sheet, coltypes_vec){
  df = read_excel(path, sheet = sheet, col_names = TRUE, col_types = coltypes_vec)
  df = as.data.frame(df)
  
  if("...1" %in% colnames(df)){
    names(df)[names(df) == '...1'] <- 'sampleID'
  }
  
  rownames(df) = df$sampleID
  df = subset(df, select = -c(sampleID))
}
```

## Loading data

```{r Load data, include = FALSE, message = FALSE, warning = FALSE}
meta = get_data(file_name, sheet_meta, coltypes_vec)
umap = get_data(file_name, sheet_umap, c("text", "numeric", "numeric"))

menst_df = get_data(menst_file, sheet_meta, c('text', 'text', 'text', 'text', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'text', 'numeric'))

nn_meta = get_data(nn_file_name, sheet_meta, append(coltypes_vec, c("text", "text")))
nn_menst_df = get_data(nn_menst_file, sheet_meta, c('text', 'text', 'text', 'text', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'text', 'numeric'))
```

## Order data

```{r UMAP, message = FALSE, warning = FALSE}
get_meta_order = function(meta){
                 meta %>% 
                 rownames_to_column("run") %>%
                 mutate_at(c('subjectID'), as.factor) %>%
                 mutate(BV = if_else(CST %in% c('I', 'II', 'III', 'V'), 'LD', 'NLD'),
                        db_type = if_else(db %in% c('cros_I', 'cros_II'), 'CROS', 'TEMP'),
                        amsel_bin = if_else(('ABV' == 1) | ('SBV' == 1), 1, 0))
}

get_temp_order = function(meta){
                  meta %>% 
                  filter(db %in% c('temp_hiseq', 'temp_pyro')) %>%
                  mutate(BV_amsel_bin = case_when((BV_status == 'ABV' | BV_status == 'SBV') ~ "1",
                                (BV_status == 'HC' & db == "carter") ~ "0",
                                (db == 'temp_pyro') ~ "no_data",
                                TRUE ~ 'no_data')) %>%
                          mutate(BV_amsel_bin = na_if(BV_amsel_bin, "no_data")) %>%
                          mutate_at(c('menst', 'BV_bin', 'BV_medication'), ~na_if(., 0))
}

all_meta = get_meta_order(meta)
all_meta = all_meta %>% 
  left_join(umap %>% rownames_to_column("run"), by = c("run")) 

nn_meta = get_meta_order(nn_meta)
# nn_meta = nn_meta %>%  
#   mutate_at(c('closest_sample', 'closest_sample_CST'), as.character) %>%
#   rename("run_orig" = "run") %>%
#   left_join(umap %>% rownames_to_column("run"), by = c("closest_sample" = "run"))
nn_meta$subjectID = str_split_fixed(nn_meta$sampleID, "_", 2)[,1]

## Temp only
temp_meta = get_temp_order(all_meta)
nn_temp_meta = get_temp_order(nn_meta)
levels(nn_temp_meta$subjectID) = levels(temp_meta$subjectID)
```

## Shannon diversity

```{r Shannon to ps}
get_bv_to_ps = function(df, x, y, fill_col, x_tit, y_tit, formula, legend_name, stat_label, colors_vec = community_colors, facet = TRUE){
  g = ggplot(df, aes(x = x, y = y)) +
  geom_point(aes(fill = fill_col), shape = 21, alpha = 0.5, size = 2) +
  stat_poly_line(color = 'black', formula = formula, fill = '#E1DEDC') +
  stat_poly_eq(use_label(stat_label), family = 'Gafata', formula = formula, size = 6) +
  scale_fill_manual(values = colors_vec, name = legend_name) +
  labs(title = '', x = x_tit, y = y_tit) +
  theme_bw(base_family = 'Gafata') +
  theme(text = element_text(size = 27),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  
  if (facet){
    g = g + facet_wrap(~df$CST, nrow = 2, scales = "free_x")
  }
  return(g)
}

health_meta = meta %>% filter(CST %in% c('I', 'II', 'III', 'V'))
s = get_bv_to_ps(meta, health_meta$mt_pseudotime, health_meta$shannon_index, health_meta$subCST, 'Pseudo-time', 'Shannon Diversity Index', y ~ poly(x, 3, raw = TRUE), 'subCST', c("R2", "p.value"), community_colors)
#y ~ poly(x, 3, raw = TRUE)
nn_s = get_bv_to_ps(nn_temp_meta, nn_temp_meta$mt_pseudotime, nn_temp_meta$shannon_index, nn_temp_meta$subCST, 'Pseudo-time', 'Shannon Diversity Index', y ~ poly(x, 3, raw = TRUE), 'subCST', c("R2", "p.value"), community_colors)
print(s)

# Save
# ggsave(filename = paste0(save_path, 'shannon_to_ps_', date, '.png'), plot = s, height = 6, width = 10, units = "in", dpi = 150)
# ggsave(filename = paste0(save_path, 'nn_shannon_to_ps_', date, '.png'), plot = nn_s, height = 6, width = 10, units = "in", dpi = 150)
```

## Shannon with study effect

```{r}
# Extract p. value
overall_p <- function(my_model) {
    f <- summary(my_model)$fstatistic
    p <- pf(f[1], f[2], f[3], lower.tail = F)
    attributes(p) <- NULL
    return(p)
}

# Create results data frame
cst_col = rep(c('I', 'II', 'III', 'V'), length(unique(meta$db)) + 1)
db_col = rep(append(unique(meta$db), "All"), each = 4)
columns = c("Dataset", "CST", "R squared", "P. value") 
db_df = data.frame(matrix(nrow = length(db_col), ncol = length(columns))) 
colnames(db_df) = columns
db_df$Dataset = db_col
db_df$CST = cst_col

db_plots_lst = list()

#Loop over db's
for (data in append(unique(meta$db), "All")){
  print(data)
  
  if (data == "All"){
    d = "All"
    db_meta = meta 
  } else {
    d = data
    db_meta = meta %>% filter(db == d)
  }
  
  for (cst_val in  unique(db_meta$CST)){
    cst_meta = db_meta %>% filter(CST == cst_val)
    if (nrow(cst_meta) > 3){
      model = lm(cst_meta$shannon_index ~ poly(cst_meta$mt_pseudotime, 3, raw = TRUE), data = cst_meta)
      model_summary <- summary(model)
      
      r_squared <- model_summary$r.squared
      p_value <- overall_p(model)
      
      db_df[which(db_df$Dataset == data & db_df$CST == cst_val), "R squared"] = r_squared
      db_df[which(db_df$Dataset == data & db_df$CST == cst_val), "P. value"] = p_value
    }
  }
}

db_df = na.omit(db_df)
db_df$Dataset[db_df$Dataset == "temp_hiseq"] <- "UAB (hiseq)"
db_df$Dataset[db_df$Dataset == "temp_pyro"] <- "UAB (pyrosequencing)"
db_df$Dataset[db_df$Dataset == "carter"] <- "PVI"
db_df$Dataset[db_df$Dataset == "cros_II"] <- "AVPVC"
db_df$Dataset[db_df$Dataset == "cros_I"] <- "VCU"
db_df$Dataset[db_df$Dataset == "srin"] <- "PHS"

write_xlsx(db_df, paste0("../data/study_effect_res.xlsx"))
```

```{r}
m1 <- lme(shannon_index ~ mt_pseudotime, random = ~1|db, data = meta)
print(anova(m1))
```



## ROC naive statistics 

```{r}
res_df = data.frame(CST = character(0), p_value = numeric(0), n = numeric(0), Method = character(0), Parameter = character(0))

roc_parameters = c('nugent', 'BV_bin', 'ph', 'clue', 'whiff', 'vag_fluid')
real_names = c('nugent' = 'Nugent', 'BV_bin' = 'Amsels test', 'ph' = 'pH', 'clue' = 'Clue cells', 'whiff' = 'Whiff test', 'vag_fluid' = 'Vaginal fluid')

for (param in roc_parameters){
  print(paste0('Working on: ', param))
  if (length(unique(meta %>% pull(param))) > 3){
    count_df = meta %>%
               filter(!is.na(!!rlang::sym(param))) %>%
               count(CST)
    
    spearman_df = meta  %>% 
                  group_by(CST) %>%
                  summarize(p_value = stats::cor.test(!!rlang::sym(param), mt_pseudotime)$p.value) %>%
                  left_join(count_df, by = 'CST') %>%
                  mutate(Method = 'Spearman', Parameter = param)
    
    res_df = res_df %>% bind_rows(spearman_df)
  }
  
  else {
    cst_to_stay = meta %>% 
                    drop_na(!!rlang::sym(param)) %>%
                    # mutate(across(c(!!rlang::sym(param)), factor)) %>%
                    group_by(CST) %>%
                    count(!!rlang::sym(param), .drop = FALSE) %>%
                    spread(!!rlang::sym(param), n) %>%
                    filter_all(all_vars(. >= 1)) %>%
                    pull(CST)
    
    count_df = meta %>%
               drop_na(!!rlang::sym(param)) %>%
               count(CST)

    wilcox_df = meta %>% 
                filter(!is.na(!!rlang::sym(param))) %>%
                filter(CST %in% cst_to_stay) %>%
                mutate(across(c(!!rlang::sym(param)), factor)) %>%
                group_by(CST) %>%
                summarize(p_value = wilcox.test(mt_pseudotime ~ !!rlang::sym(param))$p.value) %>%
                left_join(count_df, by = 'CST') %>%
                mutate(Method = 'Wilcox', Parameter = param)
    
    res_df = res_df %>% bind_rows(wilcox_df)

  }
}

res_df = res_df %>%
         mutate(Parameter = real_names[as.character(Parameter)]) %>%
         mutate(FDR = p.adjust (p_value, method = 'BH')) %>%
         rename('P. value' = 'p_value', 'CST counts' = 'n') 

write_xlsx(res_df, paste0(top_path, "/naive_statistics_res.xlsx"))

```

## Menstruation changes

```{r, fig.height = 3, fig.width = 4}
tmp = nn_menst_df %>%
            mutate(type = 'Held-out\nsamples') %>%
            bind_rows(menst_df %>% mutate(type = 'Manifold\nsamples')) %>%
            mutate_each_(funs(factor(.)), c('sum_all')) %>%
            mutate_each_(funs(as.numeric(.)), c('ps_diff_abs', 'ps_diff', 'log_ps_diff')) 

m = ggplot(data = tmp, 
           aes(x = factor(sum_all), y = (log_ps_diff)), color = 'slategray4') + 
  geom_boxplot(lwd = 0.4) +
  labs(x = '', y = "Pseudo-time\ndifferentiation (log)") +
  scale_x_discrete(labels = c('0' = 'Non-\nmenstruation', '1' = 'Menstruation\nand 2 days before')) +
  scale_y_continuous(limits = c(-10, 3)) +
  geom_signif(comparisons = list(c("0", "1")), test = "wilcox.test", test.args = list(alternative = "less", paired = FALSE, var.equal = FALSE),
              map_signif_level = TRUE, textsize = 8, tip_length = 0) + 
  facet_wrap(~factor(type, levels = c('Manifold\nsamples', 'Held-out\nsamples')), nrow = 1, ncol = 2) +
  theme_bw(base_family = 'Gafata') +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5, lineheight = 0.3),
        axis.title.y = element_text(lineheight = 0.3),
        strip.text = element_text(size = 25, lineheight = 0.3))
print(m)
ggsave(filename = paste0(save_path, 'menst_delta_14102023.png'), plot = m, dpi = 300,
       width = 3, height = 2, units = "in", limitsize = TRUE)
```

## Metabolites differences

```{r Amines}
met_df = read_excel(paste0(top_path, 'metabolites_with_ps_PCA_umap.xlsx'))
met_df$high_ps = as.factor(met_df$high_ps)
get_met_boxplot = function(df, bin_ps_col, col, colname, lab){
  p <- ggplot(data = df, aes(x = factor(bin_ps_col), y = log(col)), color = 'slategray4') +
  geom_boxplot(lwd = 0.2) +
  labs(x = 'Pseudo-time category', y = paste0(colname, ' (', expression('\u03BC'), 'M), (log)')) +
  scale_x_discrete(labels = c('0' = 'Low', '1' = 'High')) +
  scale_y_continuous(limits = c(min(log(col) - 0.5) + 1, max(log(col)) + 1)) +
  geom_signif(comparisons = list(c("0", "1")), test = "wilcox.test", test.args=list(alternative = "less"),
              map_signif_level = TRUE, textsize = 10, tip_length = 0) + #, family = 'Gafata'
  guides(fill=guide_legend(ncol=2)) +
  theme_bw(base_family = 'Gafata') +
  theme(text = element_text(size = 30),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
  return(p)
}

p1 = get_met_boxplot(met_df, met_df$high_ps, met_df$Cadaverine, 'Cadaverine', 'a')
p1 = p1 +  theme(axis.title.x = element_blank())
p2 = get_met_boxplot(met_df, met_df$high_ps, met_df$Putrescine, 'Putrescine', 'b')
p2 = p2 +  theme(axis.title.x = element_blank())
p3 = get_met_boxplot(met_df, met_df$high_ps, met_df$Spermidine, 'Spermidine', 'c')
p3 = p3 +  theme(axis.title.x = element_blank())
p4 = get_met_boxplot(met_df, met_df$high_ps, met_df$Spermine, 'Spermine', 'd')
p5 = get_met_boxplot(met_df, met_df$high_ps, met_df$Tyramine, 'Tyramine', 'e')
p6 = get_met_boxplot(met_df, met_df$high_ps, met_df$total, 'Total', 'f')


pall = plot_grid(p1, p2, p3, p4, p5, p6, nrow = 2, ncol = 3)
ggsave2(
  paste0(save_path, 'supp_figure_s6.png'),
  plot = pall,
  device = NULL,
  path = NULL,
  scale = 1,
  width = 6,
  height = 4,
  units = c("in"),
  dpi = 300)
```

```{r}
mets = ggarrange(p1, p2, p3, p4, p5, p6, nrow = 3, ncol = 3)
ggarrange(m, mets, nrow = 2, ncol = 1) %>%
ggexport(filename = paste0(save_path, 'all_17072023.png'), width = 1300, height = 1300, common.legend = TRUE, legend = "bottom")
while (!is.null(dev.list()))  dev.off()
```

## Representative pseudotime progression of three representative woman

```{r}
get_umap_subject = function(all_meta, curr_meta, subject, size, size_cst, title_size, linewidth, vjust){
  sub_df = subset(curr_meta, subjectID == subject)
  sub_df = sub_df[with(sub_df, order(time)),]
  g = ggplot() +
    geom_point(data = subset(all_meta, subjectID != subject), aes(x = fa1, y = fa2), fill = 'gray60', color = 'gray55',  
               size = size, pch = 21, alpha = 0.99) +
    geom_segment(data = sub_df, aes(x = fa1, y = fa2, xend = c(tail(fa1, n = -1), NA), yend = c(tail(fa2, n = -1), NA)),
               arrow = arrow(length = unit(0.2, "cm")),
               position = position_attractsegment(start_shave = 0.02, end_shave = 0.02),
               color = 'black', lineend = 'round', linejoin = 'mitre', linewidth = linewidth) +
    geom_point(data = sub_df, aes(x = fa1, y = fa2, fill = subCST), size = size_cst,
               pch = 21, alpha = 0.7, color = 'gray49') +
    labs(title = str_replace(subject, '_R', ''), x = "") + 
    theme_bw(base_family = 'Gafata') +
    scale_fill_manual(values = community_colors, name = 'subCST', na.value = "grey31") +
    theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        plot.title = element_text(size = title_size, family = 'Gafata', vjust = vjust, hjust =  0.03),
        legend.key = element_rect(colour = 'black', size = 1),
        plot.margin = unit(c(t = 1, r = 0.2, b = 0, l = -0.2), "cm"),
        legend.position = "none") +
        theme(text = element_text(size = 45)) 
  
  return(g)
}

get_ps_to_time = function(df, leg_pos, points_size, scale_menst_size, text_size = 30) {
  g = ggplot() +
    geom_line(data = df, aes(x = time, y = mt_pseudotime), colour = 'slategray4', alpha = 0.6, size = 0.5) +
    geom_point(data = df, aes(x = time, y = mt_pseudotime, colour = subCST), alpha = 0.6, size = points_size, pch = 19) + 
    geom_point(data = df, aes(x = time, y = -0.04, size = factor(menst)), colour = 'red', alpha = 0.9, pch = 2) +
    geom_point(data = df, aes(x = time, y = 1.05, size = factor(BV_bin)), colour = 'darkmagenta', alpha = 0.9, pch = 5) +
    geom_point(data = df, aes(x = time, y = 1.05, size = factor(BV_medication)), colour = 'darkslategray4', alpha = 0.9, pch = 5) +
    scale_colour_manual(values = community_colors) +
    scale_y_continuous(limits = c(-0.05, 1.1), breaks = seq(0.0, 1, 0.5)) +
    scale_x_continuous(limits = c(0, 75), breaks = seq(0, 70, 35)) +
    scale_size_discrete(range = scale_menst_size, guide = 'none') +
    labs(title = '', x = 'Time (days)', y = 'Pseudo-time') +
    theme_bw(base_family = 'Gafata')+
    # theme(panel.spacing.x = unit(0.15, "lines"), panel.spacing.y = unit(0.12, "lines")) +
    theme(legend.position = leg_pos, legend.spacing.y = unit(0.001, 'cm'), legend.spacing.x = unit(0.001, 'cm'),
          legend.text = element_text(margin = margin(0, 0, 0, 0))) +
    theme(text = element_text(size = 45)) + 
    theme(plot.margin = rep(unit(0, "null"), 4)) +
    guides(colour = guide_legend(override.aes = list(size = 10))) +
    theme(axis.title.y = element_text(margin = margin(0, 0, 0, 0))) +
    theme(axis.title.x = element_text(margin = margin(0, 0, 0, 0)))
  
  return(g)
}
```


```{r, warning=FALSE}
specific_subjects_lst = c('UAB103', 'UAB059', 'UAB002') 

p1 = get_umap_subject(all_meta, all_meta, 'UAB103', 0.9, 3, 45, 0.5, 0) +
    annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -1, vjust = 1.5, family = 'Gafata', size = 18) +
    labs(y = '   ') +
    theme(axis.title.y = element_text(size = 42)) +
    theme(plot.margin = unit(c(0, 0.2, 0, 0.6), "cm"),)
p2 = get_umap_subject(all_meta, all_meta, 'UAB059', 0.9, 3, 45, 0.5, 0) +
     annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -1, vjust = 1.5, family = 'Gafata', size = 18) +
     theme(axis.title.y = element_blank())
p3 = get_umap_subject(all_meta, all_meta, 'UAB002', 0.9, 3, 45, 0.5, 0) +
     annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -1, vjust = 1.5, family = 'Gafata', size = 18)  +
     theme(axis.title.y = element_blank())


specific_sub_df = subset(temp_meta, subjectID %in% specific_subjects_lst)
specific_sub_df$subjectID <- factor(specific_sub_df$subjectID, levels = specific_subjects_lst)

prog_df = get_ps_to_time(specific_sub_df, 'bottom', 3, c(1, 3), 30)
prog_df = prog_df + 
  facet_wrap(~specific_sub_df$subjectID, nrow = 1) +
  theme(strip.background = element_blank()) +
  theme(strip.text.x = element_blank()) +
  theme(plot.margin = unit(c(t = -1, r = 0.23, b = 0, l = 0), "cm")) +
  theme(axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = -15, b = 0, l = 0))) +
  theme(axis.title.x = element_text(margin = margin(t = -15, r = 0, b = 0, l = 0))) +
  theme(legend.text = element_text(size = 45)) +
  theme(panel.spacing.x = unit(0.225, "lines")) +
  guides(color = guide_legend(override.aes = list(shape = 19, size = 4), nrow = 2, title.hjust = 0.8, title.vjust = 0.5))

p_umap = plot_grid(p1, p2, p3, ncol = 3, align = "h", rel_widths = c(1, .85, .85))
pall = plot_grid(p_umap, prog_df, nrow = 2, rel_heights = c(.75, 1), align = "v")#rel_widths = c(0.6, 1))

ggsave2(
  paste0(save_path, "figure_4.png"),
  plot = pall,
  device = NULL,
  path = NULL,
  scale = 1,
  width = 8,
  height = 6,
  units = c("in"),
  dpi = 300)
```

## Bacterial changes along arms

```{r}
## Bacterial species each arm

every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}

get_barplot = function(df, title, lab){
  g = ggplot(df, aes(fill = species, y = value, x = mt_pseudotime)) + 
    geom_bar(position = "fill", stat = "identity", width = 1) +
    scale_y_continuous(expand = c(0, 0), labels = seq(0, 100, 25)) + 
    scale_x_discrete(labels = function(x) {round(as.numeric(x),2)}, breaks = every_nth(n = 5)) +
    scale_fill_manual(values = cols) +
    labs(x = 'Pseudo-time', y = 'Relative Abundance', title = paste0('', title)) +
    annotate(geom = 'text', label = lab, x = -Inf, y = Inf, hjust = 0, vjust = 1, family = 'Gafata', size = 8) +
    theme_bw(base_family = 'Gafata') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          text = element_text(size = 20))

  return(g)
}
```

```{r Bacterial species}
## Order
melt_df = get_data(paste0(top_path, "branch_melt_df_PCA_umap_02072023.xlsx"), sheet_melt_df, c("numeric", "numeric", "numeric", "text", "text"))

melt_df = melt_df[with(melt_df, order(mt_pseudotime, value)),]
melt_df$round_ps = round(melt_df$mt_pseudotime, digits = 3)
melt_df$mt_pseudotime = factor(as.character(melt_df$mt_pseudotime))
melt_df$value = round(melt_df$value, digits = 2)
melt_df$species <- gsub('_', ' ', melt_df$species)
melt_df$species <- gsub('g ', '',
                   gsub('o ', '', melt_df$species))
melt_df = melt_df %>% filter(branch %in% c('I','II','III','V'))

cols = c("Lactobacillus crispatus" = "#9d0208", 
         "Lactobacillus gasseri" = "#FFD60A", 
         "Lactobacillus iners" = "#e85d04", 
         "Lactobacillus jensenii" = "#E2D6B8", 
         "Lactobacillus Cluster 1" = "#C9184A", 
         "Lactobacillus psittaci" = "#FF758F", 
         "Gardnerella vaginalis" = "#22577a", 
         "Bifidobacterium" = "#10002b", 
         "Prevotella" = "#7b2cbf", 
         "Atopobium" = "#e0aaff", 
         "Ureaplasma" = "#d8f3dc", 
         "Anaerococcus" = "#95d5b2",
         "Megasphaera" = "#52b788", 
         "Aerococcus" = "#2d6a4f", 
         "Corynebacterium" = "#edc4b3", 
         "Streptococcus" = "#cd9777", 
         "Finegoldia" = "#774936",
         "Peptoniphilus" = "#2C2C4E",
         "Others" = "#878890")

melt_df$species <- factor(melt_df$species,
                      levels = names(cols))

get_barplot = function(df, title){
  g = ggplot(df, aes(fill = species, y = value, x = mt_pseudotime)) + 
    geom_bar(position = "fill", stat = "identity", width = 1) +
    scale_y_continuous(expand = c(0, 0), labels = seq(0, 100, 25)) +
    scale_x_discrete(labels = function(x) {round(as.numeric(x),2)}, breaks = every_nth(n = 5)) +
    scale_fill_manual(values = cols) +
    labs(x = 'Pseudo-time', y = 'Relative Abundance', title = paste0('', title)) +
    theme_bw(base_family = 'Gafata') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          text = element_text(size = 20))

  return(g)
}

## Plot
g0 = get_barplot(melt_df, 'legend') +
      theme(legend.position = 'bottom',
            legend.title = element_blank(), legend.text=element_text(size = 35)) +
      guides(fill = guide_legend(override.aes = list(size = 0.15), nrow = 7))

## Save
ggsave(paste0(save_path, 'LEGEND_bacterial_abundance_02072023.png'), plot = g0,
       width = 12, height = 8, units = 'in')
while (!is.null(dev.list()))  dev.off()

## Plots
g1 = get_barplot(subset(melt_df, branch == 'I'), 'I') +
      theme(legend.position = 'none', axis.title.x = element_blank())
# 
g2 = get_barplot(subset(melt_df, branch == 'II'), 'II') +
      theme(legend.position = 'none', axis.title.x = element_blank(), axis.title.y = element_blank())

g3 = get_barplot(subset(melt_df, branch == 'III'), 'III') +
      theme(legend.position = 'none')

g4 = get_barplot(subset(melt_df, branch == 'V'), 'V') +
      theme(legend.position = 'none')

## Save
ggarrange(g1, g2, g3, g4, nrow = 2, ncol = 2) %>%
ggexport(filename = paste0(save_path, 'bacterial_abundance_0.2_02072023.png'), width = 750, height = 620, common.legend = TRUE, legend = "bottom")
while (!is.null(dev.list()))  dev.off()
```


## UMAP each woman facet

```{r UMAP for each woman, message = FALSE, warning = FALSE}
## Plots and save
subjects_lst = mixedsort(unique(temp_meta$subjectID))
plots_lst = list()


for (i in  1:length(subjects_lst)){
  sub = subjects_lst[i]
  plot = get_umap_subject(all_meta, all_meta, sub, 0.9, 8, 60, 0.85, 0)
  plots_lst[[i]] = plot
}

pall = plot_grid(plotlist = plots_lst, nrow = 9)
ggsave2(
  paste0(save_path, "UMAP_subjectID_03072023.png"),
  plot = pall,
  device = NULL,
  path = NULL,
  scale = 1,
  width = 45,
  height = 38,
  units = c("in"),
  dpi = 300)
# p = grid.arrange(grobs = plots_lst, ncol = 9)
# ggsave(filename = paste0(save_path, 'UMAP_subjectID_03072023.png'), plot = p,
#        width = 70, height = 65, units = "in", limitsize = FALSE, dpi = 300)
```
## Pseudotime progression facet

```{r Pseudotime progression all facet}
ps_plot = get_ps_to_time(temp_meta, 'right', 2, c(0.5, 1.5),50) +
          facet_wrap(~temp_meta$subjectID, nrow = 8)
ggsave(filename = paste0(save_path, 'psToch_', date, '.png'), plot = ps_plot, height = 11, width = 9.5,   units = "in", dpi = 300)
```

## NN plots (UMAP and ps)

```{r UMAP for each woman, message = FALSE, warning = FALSE}
## Plots and save
subjects_lst = mixedsort(unique(nn_temp_meta$subjectID))
plots_lst = list()
nn_temp_meta = nn_temp_meta %>% 
  mutate_at(c('fa1', 'fa2'), as.numeric) %>%
  mutate_at(c('subCST'), as.factor)

for (i in  1:length(subjects_lst)){
  sub = subjects_lst[i]
  plot = get_umap_subject(all_meta, nn_temp_meta, sub, 0.7, 5, 150, 0.8, 0)
  plots_lst[[i]] = plot
}

pall = plot_grid(
  plotlist = plots_lst,
  scale = 0.9,
  align = 'vh',
  # hjust = -1,
  nrow = 3
)
# df, leg_pos, points_size, scale_menst_size, text_size = 30
nn_ps_plot = get_ps_to_time(nn_temp_meta, 'right', 8, c(2, 5), 150) +
           facet_wrap(~nn_temp_meta$subjectID, ncol = 4)

both = plot_grid(NULL, pall, nn_ps_plot, nrow = 1, labels = c('a', '', 'b'), label_size = 150, rel_widths = c(.05, 1, 1))

ggsave2(
  paste0(save_path, "supp_figure_s8.png"),
  plot = both,
  device = NULL,
  path = NULL,
  scale = 1,
  width = 45,
  height = 22,
  units = c("in"),
  dpi = 300)
```

## BV medications

```{r, message=FALSE, warning=FALSE}
med_meta = read_excel(paste0(top_path, "after_medication_ps_15082023.xlsx"), sheet = 'meta', col_names = TRUE) %>% select(-c("...1"))
med_meta = med_meta %>% mutate_at(c('BV_medication'), as.character) %>% 
           mutate(BV_medication = recode(BV_medication, "0" = "No", "1" = "Yes"))
label_med <- setNames(as.character(med_meta$subjectID), med_meta$group_name)

get_ps_to_time_med = function(df, leg_pos, cols, stat_label, text_size = 40) {
  g = ggplot(data = df, aes(x = time, y = mt_pseudotime)) +
    geom_point(aes(colour = BV_medication), alpha = 0.6, size = 2, pch = 19) + 
    stat_poly_line(color = 'black', formula = y ~ x, fill = '#E1DEDC', size = 0.2) +
    scale_colour_manual(values = cols, name = "Medications") +
    stat_poly_eq(use_label(stat_label), family = 'Gafata', formula = y ~ x, size = 10) +
    labs(title = '', x = 'Time (days)', y = 'Pseudo-time') +
    theme_bw(base_family = 'Gafata')+
    theme(panel.spacing.x = unit(0.15, "lines"), panel.spacing.y = unit(0.12, "lines")) +
    theme(legend.position = leg_pos, legend.spacing.y = unit(0.01, 'cm'), legend.spacing.x = unit(0.01, 'cm'),
          legend.text = element_text(size = 60, margin = margin(t = 0.2)),
          legend.title = element_text(size = 60)) +
    theme(text = element_text(size = text_size)) + 
    guides(colour = guide_legend(override.aes = list(size = 5))) 
  
  return(g)
}

m = get_ps_to_time_med(med_meta, 'bottom', c("No" = "#2E9B8F", "Yes" = "#E76F51"), c("eq")) +
    facet_wrap(~med_meta$group_name, nrow = 3, scales = "free", labeller = as_labeller(label_med)) +
    theme(legend.position = 'bottom')
    

ggsave(filename = paste0(save_path, 'supp_figure_s7.png'), plot = m, height = 6, width = 8, units = "in", dpi = 300)

```

## Heatmap

```{r}
db_colors = c('UAB_pyro' = '#F7C59F', 'UAB_hiseq' = '#EFEFD0', 'VCU' = '#89B1D0', 'AVPVC' = '#1A659E', 'PVI' = '#FA5C23', 'PHS' = '#00365F')
db_colors = c('UAB_pyro' = '#F7C59F', 'UAB_hiseq' = '#EFEFD0', 'VCU' = '#89B1D0', 'AVPVC' = '#1A659E', 'PVI' = '#FA5C23', 'PHS' = '#00365F')
plot_heatmap <- function(df, vec_cst, vec_score, n = 25) {
  # Order for colsums
  df = df %>%
    column_to_rownames("sample_id")
  
  # Create Others column
  df <- df[, order(colSums(-df))]
  others_vec = rowSums(df[n:ncol(df)])
  df = df[, 1:n]
  df$Others <- others_vec
  
  # Merge metadata
  all_df = merge(df, h_meta, by = 'row.names')
  all_df = all_df %>%
    column_to_rownames(var = 'Row.names') %>%
    arrange(subCST, score, db)
  
  plot_mat = as.matrix(t(all_df %>% select(-c('subCST', 'CST', 'score', 'db'))))
  
  # Plot
  col_fun = colorRamp2(seq(0, 100, length = 6), c("#F1F3F3", "#DDFFF7", "#93E1D8", "#FFA69E", "#AA4465", "#462255"))
  column_ha = HeatmapAnnotation(subCST = all_df$subCST,
                                  Score = anno_barplot(all_df$score, baseline = 0, height = unit(0.7, "cm"),
                                                       gp = gpar(col = "#462255"), axis_param = list(at = c(0, 1))),
                                  Dataset = all_df$db,
                                  col = list(subCST = community_colors, Dataset = db_colors),
                                  annotation_legend_param = list(subCST = list(nrow = 3, 
                                                                               legend_width = unit(20, "cm"),
                                                                               legend_height = unit(5, "cm"),
                                                                               grid_width = unit(5, "mm"),
                                                                               legend_gp = gpar(fontsize = 15),                                                                                                labels_gp = gpar(fontsize = 13),
                                                                               title_gp  = gpar(fontsize = 15)),
                                                                 Dataset = list(nrow = 3, 
                                                                                legend_width = unit(20, "cm"),
                                                                                legend_height = unit(5, "cm"),
                                                                                grid_width = unit(5, "mm"),
                                                                                legend_gp = gpar(fontsize = 15), 
                                                                                labels_gp = gpar(fontsize = 13),
                                                                                title_gp  = gpar(fontsize = 15))),
                                                                 
                                  simple_anno_size = unit(0.8, "cm"),
                                  annotation_name_gp = gpar(fontsize = 20, fontface = "italic"))
  h = Heatmap(plot_mat, name = 'Abundance', col = col_fun, 
              show_column_names = FALSE,
              top_annotation = column_ha, 
              cluster_rows = FALSE, cluster_columns = FALSE,
              row_names_gp = gpar(fontsize = 18),
              row_names_max_width = unit(12, "cm"),
              heatmap_legend_param = list(legend_direction = "horizontal", legend_width = unit(60, "mm"), legend_height = unit(50, "mm"), labels_gp  = gpar(fontsize = 12), title_gp  = gpar(fontsize = 14)))
  return(h)
  }
```

```{r, messege = F}
df = read_excel("C:/Users/morts/Documents/TAU/Manifold_Project/Data/Serrano_Ravel13_Ravel11_Carter22/all_df_26062023.xlsx", sheet = 'abundance', col_names = TRUE)
df = df %>% 
  rename_with(~ gsub("_", " ", .x, fixed = TRUE)) %>%
  rename(sample_id = 1) %>%
  rename_with(~ gsub("g ", "", .x, fixed = TRUE)) %>%
  rename_with(~ gsub("o ", "", .x, fixed = TRUE)) 
```
```{r, warning = FALSE, message=FALSE}
meta = read_excel("C:/Users/morts/Documents/TAU/Manifold_Project/Data/Serrano_Ravel13_Ravel11_Carter22/all_df_26062023.xlsx", sheet = 'meta', col_names = TRUE)
h_meta = meta %>%
  rename(sample_id = 1) %>%
  column_to_rownames("sample_id") %>%
  select(c('subCST', 'CST', 'score', 'db')) %>%
  mutate(db = case_when(db == 'temp_hiseq' ~ 'UAB_hiseq',
                        db == 'temp_pyro' ~ 'UAB_pyro',
                        db == 'carter' ~ 'PVI',
                        db == 'cros_II' ~ 'AVPVC',
                        db == 'cros_I' ~ 'VCU',
                        db == 'srin' ~ 'PHS'))
h_meta = h_meta[match(df$sample_id, rownames(h_meta)),]
```

```{r}
h = plot_heatmap(df, h_meta$subCST, h_meta$score)# get_h(h_dft, cst_df$subCST, cst_df$score)
png(paste(save_path, 'all_df_02072023.png', sep = ''), 
    width = 720, height = 800)
draw(h, padding = unit(c(2, 2, 2, 20), "mm"),
     heatmap_legend_side = "bottom", annotation_legend_side = "bottom", gap = unit(1, "cm"))#, ht_gap = unit(7, "mm"))
dev.off()
```

```{r, fig.height = 4, fig.width=8}
cst_colors =  c('I' = '#9d0208', 'II' = '#FFD60A', 'III' = '#e85d04','IV-A' = '#22577a', 'IV-B' = '#5390D9', 'IV-C' = '#2E214F', 'V' = '#E2D6B8')

df <- count(meta, CST)
v = ggplot(data = meta, aes(x = CST, y = shannon_index, fill = CST)) +
    geom_boxplot() +
    # stat_summary(fun.data = stat_box_data, geom = "text", hjust = 0.5, vjust = 0.9, size = 8) + 
    geom_text(data = df, aes(y = 5.2, label = n), size = 20) +
    scale_fill_manual(values = cst_colors, name = 'CST', na.value = "grey31") +
    theme_bw(base_family = 'Gafata') +
    labs(x = 'subCST', y = 'Shannon diversity index') +
    theme(text = element_text(size = 65)) +
    scale_y_continuous(limits = c(0, 5.5)) +
    theme(legend.position = 'bottom') +
    guides(fill=guide_legend(nrow=2))
ggsave(paste0(save_path, 'supp_figure_s1_b.png'), height = 6, width = 8, units = "in", dpi = 300)
print(v)
```

```{r}
plot_grid(h, v, nrow = 2, ncol = 1, labels = c('A', 'B'), heights = c(.75, .25))
ggsave2(
  paste0(top_path, 'h.png'),
  width = NA,
  height = NA,
  units = c("in", "cm", "mm"),
  dpi = 300,
  limitsize = TRUE
)
```



