---
title: "All plots- manifold project"
author: "Borenstein Lab"
date: "March 2023"
output:
  html_document:
    css: custom_notebook_formatting.css
    toc: true
    toc_depth: 3
    df_print: paged
    code_folding: hide
---

```{r setup, include = FALSE}
rm(list=ls())

library(vegan)
library(ggarchery)
library(data.table)
library(ggpubr)
library(ggpmisc)
library(scales)
library(readxl)
library(reshape2)
library(dplyr)
library(tidyr)
library(stringr)
library(writexl)
library(ggpubr)
library(gridExtra)
library(grid)
library(RColorBrewer)
library(xlsx)
library(tibble)
library(umap)
library(tseries)
library(seastests)
library(extrafont)
library(showtext)
font_add_google("Gafata")
showtext_auto()
```

## Variables

```{r variables}
top_path = '../../Data/Serrano_Ravel13_Ravel11_Carter22/ps_res/'

file_name = paste0(top_path, 'ps_res_18062023_BC_X.xlsx')

sheet_abun = 'abundance'
sheet_meta = 'meta'
sheet_umap = 'umap'

roc_file = paste0(top_path, 'roc_res_17052023.xlsx')
menst_file = paste0(top_path, 'menst_statistics_res.xlsx')

bactria_file_name = paste0(top_path, 'branch_melt_df_16052023.xlsx')
sheet_melt_df = 'melt_branch_df'

long_bv_diff = paste0(top_path, "10days_without_bv_diff_04062023.xlsx")
sheet_long = 'bv'

nn_file_name = paste0(top_path, 'nn_res_30052023.xlsx')
nn_roc_file = paste0(top_path, 'bv_redict_ps_31052023.xlsx')
nn_menst_file = paste0(top_path, 'nn_menst_statistics_res.xlsx')


save_path = paste0(top_path, 'figures/')

date = '30052023'
```

## Colors pallettes

```{r}
community_colors = c('I-A' = '#9d0208', 'I-B' = '#dc2f02', 'II' = '#FFD60A', 'III-A' = '#e85d04','III-B' = '#faa307','IV-A' = '#22577a', 'IV-B' = '#5390D9', 'IV-C0' = '#2E214F', 'IV-C1' = '#63479B','IV-C2' = '#8962C7', 'IV-C3' = '#AB7DDA', 'IV-C4' = '#DEC9E9', 'V' = '#E2D6B8')

db_colors = c('temp_pyro' = '#006d77', 'temp_hiseq' = '#83c5be', 'cros_I' = '#edf6f9', 'cros_II' = '#ffddd2', 'carter' = '#E6744A')

coltypes_vec = c("text", "text", "text", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", 
                 "text", "text", "numeric", "text", "numeric", "numeric", "numeric", "numeric", "text",
                 "text", "numeric", "numeric", "numeric", "text", "numeric", "numeric", "numeric", "numeric", 
                 "text", "numeric", "text", "numeric", "numeric")
```

```{r}
get_data = function(path, sheet, coltypes_vec){
  df = read_excel(path, sheet = sheet, col_names = TRUE, col_types = coltypes_vec)
  df = as.data.frame(df)
  
  if("...1" %in% colnames(df)){
    names(df)[names(df) == '...1'] <- 'sampleID'
  }
  
  rownames(df) = df$sampleID
  df = subset(df, select = -c(sampleID))
}
```

## Loading data

```{r Load data, include = FALSE, message = FALSE, warning = FALSE}
meta = get_data(file_name, sheet_meta, coltypes_vec)
umap = get_data(file_name, sheet_umap, c("text", "numeric", "numeric"))
bacteria_df = get_data(bactria_file_name, sheet_melt_df, c("numeric", "numeric", "numeric", "text", "text"))
menst_df = get_data(menst_file, sheet_meta, c('text', 'text', 'text', 'text', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'text', 'numeric','text', 'text', 'text'))
long_bv_df = df = read_excel(long_bv_diff, sheet = sheet_long, col_names = TRUE)

nn_meta = get_data(nn_file_name, sheet_meta, append(coltypes_vec, c("text", "numeric", "numeric", "text")))
nn_menst_df = get_data(nn_menst_file, sheet_meta, c('text', 'text', 'text', 'text', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'numeric', 'text', 'numeric','text', 'text', 'text'))
# nn_bv_roc = get_data(nn_roc_file)
```

## Order data

```{r UMAP, message = FALSE, warning = FALSE}
get_meta_order = function(meta){
                 meta %>% 
                 rownames_to_column("run") %>%
                 mutate_at(c('subjectID'), as.factor) %>%
                 mutate(BV = if_else(CST %in% c('I', 'II', 'III', 'V'), 'LD', 'NLD'),
                        db_type = if_else(db %in% c('cros_I', 'cros_II'), 'CROS', 'TEMP'),
                        amsel_bin = if_else(('ABV' == 1) | ('SBV' == 1), 1, 0))
}

get_temp_order = function(meta){
                  meta %>% 
                  filter(db %in% c('temp_hiseq', 'temp_pyro')) %>%
                  mutate(BV_amsel_bin = case_when((BV_status == 'ABV' | BV_status == 'SBV') ~ "1",
                                (BV_status == 'HC' & db == "carter") ~ "0",
                                (db == 'temp_pyro') ~ "no_data",
                                TRUE ~ 'no_data')) %>%
                          mutate(BV_amsel_bin = na_if(BV_amsel_bin, "no_data")) %>%
                          mutate_at(c('menst', 'amsel_bin', 'BV_medication'), ~na_if(., 0))
}

all_meta = get_meta_order(meta)
all_meta = all_meta %>% 
  left_join(umap %>% rownames_to_column("run"), by = c("run")) 
nn_meta = get_meta_order(nn_meta)
nn_meta = nn_meta %>%  
  rename("run_orig" = "run") %>%
  left_join(umap %>% rownames_to_column("run"), by = c("closest_sample" = "run"))
nn_meta$subjectID = str_split_fixed(nn_meta$sampleID, "_", 2)[,1]

## Temp only
temp_meta = get_temp_order(all_meta)
nn_temp_meta = get_temp_order(nn_meta)
levels(nn_temp_meta$subjectID) = levels(temp_meta$subjectID)
```

## Functions

```{r All plot functions}

## UMAP each subject

get_umap_subject = function(all_meta, curr_meta, subject, size, size_cst, title_size, linewidth, vjust){
  sub_df = subset(curr_meta, subjectID == subject)
  sub_df = sub_df[with(sub_df, order(time)),]
  g = ggplot() +
    geom_point(data = subset(all_meta, subjectID != subject), aes(x = fa1, y = fa2), fill = 'gray60', color = 'gray55',  
               size = size, pch = 21, alpha = 0.99) +
    geom_segment(data = sub_df, aes(x = fa1, y = fa2, xend = c(tail(fa1, n = -1), NA), yend = c(tail(fa2, n = -1), NA)),
               arrow = arrow(length = unit(0.2, "cm")),
               position = position_attractsegment(start_shave = 0.02, end_shave = 0.02),
               color = 'black', lineend = 'round', linejoin = 'mitre', linewidth = linewidth) +
    geom_point(data = sub_df, aes(x = fa1, y = fa2, fill = subCST), size = size_cst,
               pch = 21, alpha = 0.7, color = 'gray49') +
    labs(title = str_replace(subject, '_R', ''), x = NULL, y = NULL) + 
    theme_bw(base_family = 'Gafata') +
    scale_fill_manual(values = community_colors, name = 'subCST', na.value = "grey31") +
    theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(size = title_size, family = 'Gafata', vjust = vjust, hjust =  0.03),
        legend.position = "none",
        legend.key = element_rect(colour = 'black', size = 1),
        plot.margin = rep(unit(0, "null"), 4),
        panel.spacing = unit(0,"null"),)
  
  return(g)
}

## Ps to time

get_ps_to_time = function(df, leg_pos, points_size, scale_menst_size, text_size = 30) {
  g = ggplot() +
    geom_line(data = df, aes(x = time, y = mt_pseudotime), colour = 'slategray4', alpha = 0.6, size = 0.5) +
    geom_point(data = df, aes(x = time, y = mt_pseudotime, colour = subCST), alpha = 0.6, size = points_size, pch = 19) + 
    geom_point(data = df, aes(x = time, y = -0.04, size = factor(menst)), colour = 'red', alpha = 0.9, pch = 2) +
    geom_point(data = df, aes(x = time, y = 1.05, size = factor(amsel_bin)), colour = 'darkmagenta', alpha = 0.9, pch = 5) +
    geom_point(data = df, aes(x = time, y = 1.05, size = factor(BV_medication)), colour = 'darkslategray4', alpha = 0.9, pch = 5) +
    scale_colour_manual(values = community_colors) +
    scale_y_continuous(limits = c(-0.05, 1.1), breaks = seq(0.0, 1, 0.5)) +
    scale_x_continuous(limits = c(0, 75), breaks = seq(0, 70, 35)) +
    scale_size_discrete(range = scale_menst_size, guide = 'none') +
    labs(title = '', x = 'Time (days)', y = 'Pseudo-time') +
    theme_bw(base_family = 'Gafata')+
    theme(panel.spacing.x = unit(0.15, "lines"), panel.spacing.y = unit(0.12, "lines")) +
    theme(strip.text = element_text(size = 60, lineheight = 0.3),
        strip.background = element_rect(fill = "antiquewhite2", colour = "black"), margin = margin(0, 0, 0, 0)) + 
    theme(legend.position = leg_pos, legend.spacing.y = unit(0.01, 'cm'), legend.spacing.x = unit(0.01, 'cm'),
          legend.text = element_text(margin = margin(t = 0.2))) +
    theme(text = element_text(size = text_size)) + 
    guides(colour = guide_legend(override.aes = list(size = 10)))
  
  return(g)
}


## Shannon

get_bv_to_ps = function(df, x, y, x_tit, y_tit, formula, x_lab){
  g = ggplot(df, aes(x = x, y = y)) +
  geom_point(aes(fill = subCST), shape = 21, alpha = 0.5, size = 2) +
  geom_smooth(method = "glm", color = 'black', family=gaussian(link="log")) +

  # stat_poly_line(color = 'black', formula = formula) +
  stat_poly_eq(family = 'Gafata', formula = formula, size = 5) +
  scale_fill_manual(values = community_colors, name = 'subCST') +
  labs(title = '', x = x_tit, y = y_tit) +
  facet_wrap(~df$CST, nrow = 2, scales = "free_x") +
  theme_bw(base_family = 'Gafata') +
  theme(text = element_text(size = 20)) 
  return(g)
}

## Bacterial species each arm

every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}

get_barplot = function(df, title, lab){
  g = ggplot(df, aes(fill = species, y = value, x = mt_pseudotime)) + 
    geom_bar(position = "fill", stat = "identity", width = 1) +
    scale_y_continuous(expand = c(0, 0), labels = seq(0, 100, 25)) + 
    scale_x_discrete(labels = function(x) {round(as.numeric(x),2)}, breaks = every_nth(n = 5)) +
    scale_fill_manual(values = cols) +
    labs(x = 'Pseudo-time', y = 'Relative Abundance', title = paste0('', title)) +
    annotate(geom = 'text', label = lab, x = -Inf, y = Inf, hjust = 0, vjust = 1, family = 'Gafata', size = 8) +
    theme_bw(base_family = 'Gafata') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          text = element_text(size = 20))

  return(g)
}
```

## UMAP each woman

```{r UMAP for each woman, message = FALSE, warning = FALSE}
## Plots and save
subjects_lst = unique(nn_temp_meta$subjectID)
plots_lst = list()


for (i in  1:length(subjects_lst)){
  sub = subjects_lst[i]
  plot = get_umap_subject(all_meta, nn_temp_meta, sub, 0.7, 5, 60, 0.8, -2)
  plots_lst[[i]] = plot
}

nn_p = grid.arrange(grobs = plots_lst, ncol = 4)
nn_ps_plot = get_ps_to_time(nn_temp_meta, 'right', 4, c(2, 4), 60) +
          facet_wrap(~nn_temp_meta$subjectID, nrow = 4)

nn_pall = grid.arrange(arrangeGrob(nn_p, nn_ps_plot, ncol = 2, widths = c(1.5, 1)))
# nn_pall = plot_grid(nn_p, nn_ps_plot, labels = c('A', 'B'))
ggsave(filename = paste0(save_path, 'nn_UMAP_subjectID_30052023.png'), plot = nn_pall,
       width = 25, height = 10, units = "in", limitsize = FALSE)

# p = grid.arrange(grobs = plots_lst, ncol = 9)
# ggsave(filename = paste0(save_path, 'nn_UMAP_subjectID_30052023.png'), plot = p,
#        width = 50, height = 55, units = "in", limitsize = FALSE)
```

## Representative pseudotime progression of four representative woman

```{r}
labels = c('A', 'B', 'C', 'D')
specific_subjects_lst = c('UAB103', 'UAB059', 'UAB083') #UAB012

p1 = get_umap_subject('UAB103', 0.9, 3, 30, 0.5, 0) +
    labs(y = 'UMAP') +
    annotate(geom = 'text', label = 'A', x = -Inf, y = Inf, hjust = -0.2, vjust = 1.5, family = 'Gafata', size = 7) +
    theme(axis.title = element_text(size = 42, color = 'white'))
# p2 = get_umap_subject('UAB012', 0.9, 3, 30, 0.5, 0) +
#      annotate(geom = 'text', label = 'B', x = -Inf, y = Inf, hjust = -0.2, vjust = 1.5, family = 'Gafata', size = 7) 
p2 = get_umap_subject('UAB059', 0.9, 3, 30, 0.5, 0) +
     annotate(geom = 'text', label = 'B', x = -Inf, y = Inf, hjust = -0.2, vjust = 1.5, family = 'Gafata', size = 7) 
p3 = get_umap_subject('UAB083', 0.9, 3, 30, 0.5, 0) +
     annotate(geom = 'text', label = 'C', x = -Inf, y = Inf, hjust = -0.2, vjust = 1.5, family = 'Gafata', size = 7) 


specific_sub_df = subset(temp_meta, subjectID %in% specific_subjects_lst)
specific_sub_df$subjectID <- factor(specific_sub_df$subjectID, levels = specific_subjects_lst)
specific_sub_df$lab <- with(specific_sub_df, ifelse(subjectID == 'UAB103', 'D',
                                             ifelse(subjectID == 'UAB059', 'E', 
                                             ifelse(subjectID == 'UAB083', 'F', 'G'))))

prog_df = get_ps_to_time(specific_sub_df, 'bottom', 3, c(1, 3), 30)
prog_df = prog_df + 
  geom_text(data = specific_sub_df, label = specific_sub_df$lab, x = -Inf, y = Inf, hjust = -0.2, vjust = 1.5, family = 'Gafata', size = 7) +
  # geom_text(data = specific_sub_df, aes(x = 10, y = 0.4, label = lab))
  facet_wrap(~specific_sub_df$subjectID, nrow = 1) +
  # geom_text(aes(x, y, label=lab),
  #       data = data.frame(x = 6, y = 0.5, lab = c("(d)","(e)","(f)"), subjectID = specific_subjects_lst), vjust = 1) +
  theme(strip.background = element_blank(),
  axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
  strip.text.x = element_blank(),
  plot.margin = rep(unit(0, "null"), 4),
  legend.spacing.y = unit(0, "lines"),
  legend.spacing.x = unit(1.0, 'cm')) +
  guides(fill = guide_legend(override.aes = list(shape = 21, size = 7), ncol = 2, title.hjust = 2, title.vjust = 1))

pall = grid.arrange(arrangeGrob(p1, p2, p3, ncol = 3, widths = c(1.25, 1, 1)), prog_df, heights = c(1.5/4, 2.5/4), nrow = 2)

ggsave(filename = paste0(save_path, 'UMAP_subjectID_example_18052023.png'), plot = pall,
       width = 10, height = 8, units = "in", limitsize = TRUE)
```

## Pseudotime progression facet

```{r Pseudotime progression all facet}
ps_plot = get_ps_to_time(temp_meta, 'right', 2, c(0.5, 1.5),30) +
          facet_wrap(~temp_meta$subjectID, nrow = 8)
## Save
ggsave(filename = paste0(save_path, 'psToch_', date, '.png'), plot = ps_plot, height = 11, width = 9.5,   units = "in", dpi = 150)
```


## NN plot (UMAP and ps)

```{r UMAP for each woman, message = FALSE, warning = FALSE}
## Plots and save
subjects_lst = unique(nn_temp_meta$subjectID)
plots_lst = list()


for (i in  1:length(subjects_lst)){
  sub = subjects_lst[i]
  plot = get_umap_subject(all_meta, nn_temp_meta, sub, 0.7, 5, 60, 0.8, -7)
  plots_lst[[i]] = plot
}

nn_p = grid.arrange(grobs = plots_lst, ncol = 4)

get_ps_to_time = function(df, leg_pos, points_size, scale_menst_size, text_size = 30) {
  g = ggplot() +
    geom_line(data = df, aes(x = time, y = mt_pseudotime), colour = 'slategray4', alpha = 0.6, size = 0.5) +
    geom_point(data = df, aes(x = time, y = mt_pseudotime, colour = subCST), alpha = 0.6, size = points_size, pch = 19) + 
    geom_point(data = df, aes(x = time, y = -0.04, size = factor(menst)), colour = 'red', alpha = 0.9, pch = 2) +
    geom_point(data = df, aes(x = time, y = 1.05, size = factor(amsel_bin)), colour = 'darkmagenta', alpha = 0.9, pch = 5) +
    geom_point(data = df, aes(x = time, y = 1.05, size = factor(BV_medication)), colour = 'darkslategray4', alpha = 0.9, pch = 5) +
    scale_colour_manual(values = community_colors) +
    scale_y_continuous(limits = c(-0.05, 1.1), breaks = seq(0.0, 1, 0.5)) +
    scale_x_continuous(limits = c(0, 75), breaks = seq(0, 70, 35)) +
    scale_size_discrete(range = scale_menst_size, guide = 'none') +
    labs(title = '', x = 'Time (days)', y = 'Pseudo-time') +
    theme_bw(base_family = 'Gafata')+
    theme(panel.spacing.x = unit(0.15, "lines"), panel.spacing.y = unit(0.12, "lines")) +
    theme(strip.text = element_text(size = 60, lineheight = 0.3),
        strip.background = element_rect(fill = "antiquewhite2", colour = "black")) + 
    theme(legend.position = leg_pos, legend.spacing.y = unit(0.01, 'cm'), legend.spacing.x = unit(0.01, 'cm'),
          legend.text = element_text(margin = margin(t = 0.2))) +
    theme(text = element_text(size = text_size)) + 
    guides(colour = guide_legend(override.aes = list(size = 10)))
  
  return(g)
}


nn_ps_plot = get_ps_to_time(nn_temp_meta, 'right', 6, c(2, 5), 60) +
          facet_wrap(~nn_temp_meta$subjectID, nrow = 4)

nn_pall = grid.arrange(arrangeGrob(nn_p, nn_ps_plot, ncol = 2, widths = c(1.5, 1)))
# nn_pall = plot_grid(nn_p, nn_ps_plot, labels = c('A', 'B'))
ggsave(filename = paste0(save_path, 'nn_UMAP_subjectID_05062023.png'), plot = nn_pall,
       width = 35, height = 25, units = "in", limitsize = FALSE)

# p = grid.arrange(grobs = plots_lst, ncol = 9)
# ggsave(filename = paste0(save_path, 'nn_UMAP_subjectID_30052023.png'), plot = p,
#        width = 50, height = 55, units = "in", limitsize = FALSE)
```

## Autocorrelation

```{r}
all_acf_df = data.frame()
for (subject in unique(temp_meta$subjectID)){
  subject_df = temp_meta %>% filter(subjectID == subject) %>% drop_na(mt_pseudotime)
  acf_ser <- acf(subject_df$mt_pseudotime, pl = FALSE, lag.max = 20)
  acf_df <- data.frame(lag = acf_ser$lag[,,1], acf =  acf_ser$acf[,,1])
  acf_df['subjectID'] <- subject
  
  all_acf_df <- rbind(all_acf_df, acf_df)
  
}

a = ggplot(data = all_acf_df) +
    # geom_line(data = df, aes(x = time, y = mt_pseudotime), colour = 'slategray4', alpha = 0.6, size = 0.5) +
    geom_point(aes(x = lag, y = acf), alpha = 0.6, size = 2.5, pch = 19) + 
    scale_colour_manual(values = community_colors) +
    # scale_y_continuous(limits = c(-0.05, 1.1), breaks = seq(0.0, 1, 0.5)) +
    # scale_x_continuous(limits = c(0, 75), breaks = seq(0, 70, 35)) +
    labs(title = '', x = 'Lags', y = 'Autocorrelation') +
    facet_wrap(~ subjectID) +
    theme_bw(base_family = 'Gafata')+
    theme(panel.spacing.x = unit(0.15, "lines"), panel.spacing.y = unit(0.12, "lines")) +
    theme(strip.text = element_text(face = "bold", size = 20),
        strip.background = element_rect(fill = "slategray3", colour = "black", size = 0.5)) +
    theme(legend.position = 'none') +
    theme(text = element_text(size = 20))

ggsave(filename = paste0(save_path, 'Autocorrelation_10052023.png'), plot = a, height = 11, width = 9.5,   units = "in", dpi = 150)
  
subject = 'UAB040'
subject_df = temp_meta %>% filter(subjectID == subject) %>% drop_na(mt_pseudotime)
kw(subject_df$mt_pseudotime, freq = 10, diff = T, residuals = F, autoarima = T)
isSeasonal(subject_df$mt_pseudotime, test = "combined", freq = 30)

```


## Clusters of pseudotime progression

```{r Pseudotime progression clustering, message = FALSE, warning = FALSE}
clust_meta = merge(meta, ps_clust[,c('subjectID', 'k')], by = 'subjectID')
clust_meta$k = clust_meta$k + 1
subject_colors <- colorRampPalette(brewer.pal(11, "BrBG"))(length(unique(clust_meta$subjectID)))

c = ggplot(data = clust_meta) +
    geom_line(aes(x = time, y = mt_pseudotime, colour = subjectID), alpha = 0.6, size = 0.8) +
    geom_point(aes(x = time, y = mt_pseudotime, size = menst, colour = subjectID), alpha = 0.9, pch = 2) +
    facet_wrap(~ k) +
    theme(legend.position = 'none') +
    scale_colour_manual(values = subject_colors) +
    scale_y_continuous(limits = c(-0.05, 1.1), breaks = seq(0.0, 1, 0.5)) +
    scale_x_continuous(limits = c(0, 75), breaks = seq(0, 70, 35)) +
    scale_size_continuous(range = c(1, 1), guide = 'none') +
    labs(title = '', x = 'Time (days)', y = 'Pseudo-time') +
    theme_bw(base_family = 'Gafata')+
    theme(panel.spacing.x = unit(0.15, "lines"), panel.spacing.y = unit(0.12, "lines")) +
    theme(strip.text = element_text(face = "bold", size = 18),
        strip.background = element_rect(fill = "seashell2", colour = "black", size = 0.5)) +
    theme(legend.position = 'none', legend.spacing.y = unit(0.01, 'cm'), legend.spacing.x = unit(0.01, 'cm'),
          legend.text = element_text(margin = margin(t = 0.2))) +
    theme(text = element_text(size = 18)) + 
    guides(colour = guide_legend(override.aes = list(size = 3)))
print(c)

ggsave(filename = paste0(save_path, 'clustered_ps_trajectories_02042023.png'), plot = c, height = 5, width = 6,   units = "in", dpi = 150)
```

## Menstruation changes

```{r, fig.height = 3, fig.width = 4}
# wilcox.test(x = menst_df$ps_diff_abs[menst_df$change_menst_status==TRUE],
#             y = menst_df$ps_diff_abs[menst_df$change_menst_status==FALSE],
#             mu=0, alt="two.sided", conf.int=T, conf.level=0.95,
#             paired=FALSE, exact=T, correct=T)

# menst_group = subset(menst_df$ps_diff_abs, change_menst_status == TRUE)
# wilcox.test(menst_df$ps_diff_abs[menst_df$change_menst_status==TRUE] ~ menst_df$ps_diff_abs[menst_df$change_menst_status==FALSE]) 

stat_box_data <- function(y, upper_limit = max(iris$Sepal.Length) * 1.15) {
  return( 
    data.frame(
      y = 0.95 * upper_limit,
      label = paste('count =', length(y), '\n',
                    'mean =', round(mean(y), 1), '\n')
    )
  )
}


tmp = nn_menst_df %>%
            mutate(type = 'Manifold\nsamples') %>%
            bind_rows(menst_df %>% mutate(type = 'Excluded\nsamples')) %>%
            drop_na(ps_diff_abs) %>%
            mutate_each_(funs(factor(.)), c('sum_all', 'sum_without', 'sum_closer_without', 'sum_closer')) %>%
            mutate_each_(funs(as.numeric(.)), c('ps_diff_abs', 'ps_diff')) 

m = ggplot(data = tmp, 
           aes(x = factor(sum_closer), y = ps_diff), color = 'slategray4') +
  geom_boxplot(lwd = 0.4) +
  labs(x = '', y = "Pseudo-time\ndifferentiation") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_discrete(labels = c('0' = 'Non-\nmenstruation', '1' = 'Menstruation\n +- 2 days')) +
  geom_signif(comparisons = list(c("0", "1")), test = "wilcox.test", test.args=list(alternative = "less", paired = FALSE, var.equal = FALSE),
              map_signif_level = TRUE, textsize = 20, tip_length = 0) + #, family = 'Gafata'
  stat_summary(fun.data = stat_box_data, geom = "text", fun.y = mean,
                  family = 'Gafata', hjust = 0.5, vjust = 0.9, size = 5) +
  facet_wrap(~factor(type, levels=c('Manifold\nsamples', 'Excluded\nsamples')), nrow = 1, ncol = 2) +
  theme_bw(base_family = 'Gafata') +
  theme(text = element_text(size = 50),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5, lineheight = 0.3),
        axis.title.y = element_text(lineheight = 0.3),
        strip.text = element_text(size = 50, lineheight = 0.3))
print(m)
ggsave(filename = paste0(save_path, 'menst_delta_30052023.png'), plot = m,
       width = 8, height = 4.5, units = "in", limitsize = TRUE)
```

## 10 days difference between BV and not BV

```{r}
tmp = long_bv_df %>%
  filter(!subjectID %in% (c("UAB021"))) %>%
  filter(min >= 0.899)
      
d = ggplot(data = tmp, 
           aes(x = factor(late), y = std), color = 'slategray4') +
  geom_boxplot(lwd = 0.4) +
  # labs(x = '', y = "Pseudo-time\ndifferentiation") +
  # scale_y_continuous(limits = c(0, 1)) +
  # scale_x_discrete(labels = c('0' = 'no BV', '1' = 'BV')) +
  geom_signif(comparisons = list(c("0", "1")), test = "wilcox.test", test.args=list(alternative = "less", paired = FALSE, var.equal = FALSE),
              map_signif_level = TRUE, textsize = 20, tip_length = 0) + #, family = 'Gafata'
  # stat_summary(fun.data = stat_box_data, geom = "text", fun.y = mean,
  #                 family = 'Gafata', hjust = 0.5, vjust = 0.9, size = 5) +
  # facet_wrap(~factor(type, levels=c('Manifold\nsamples', 'Excluded\nsamples')), nrow = 1, ncol = 2) +
  theme_bw(base_family = 'Gafata') +
  theme(text = element_text(size = 50),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5, lineheight = 0.3),
        axis.title.y = element_text(lineheight = 0.3),
        strip.text = element_text(size = 50, lineheight = 0.3))
# print(m)
# ggsave(filename = paste0(save_path, 'menst_delta_30052023.png'), plot = m,
#        width = 8, height = 4.5, units = "in", limitsize = TRUE)
```


## Shannon diversity to pseudotime by branch

```{r Shannon to ps}
## Plots
get_bv_to_ps = function(df, x, y, x_tit, y_tit, formula, x_lab, facet = TRUE){
  g = ggplot(df, aes(x = x, y = y)) +
  geom_point(aes(fill = subCST), shape = 21, alpha = 0.5, size = 2) +
  stat_poly_line(color = 'black', formula = formula) +
  stat_poly_eq(family = 'Gafata', formula = formula, size = 6) +
  scale_fill_manual(values = community_colors, name = 'subCST') +
  labs(title = '', x = x_tit, y = y_tit) +
  theme_bw(base_family = 'Gafata') +
  theme(text = element_text(size = 27),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  
  if (facet){
    g = g + facet_wrap(~df$CST, nrow = 2, scales = "free_x")
  }
  return(g)
}

s = get_bv_to_ps(temp_meta, temp_meta$mt_pseudotime, temp_meta$shannon_index, 'Pseudo-time', 'Shannon Diversity Index', y ~ poly(x, 3, raw = TRUE) , 4.2)

nn_s = get_bv_to_ps(nn_temp_meta, nn_temp_meta$mt_pseudotime, nn_temp_meta$shannon_index, 'Pseudo-time', 'Shannon Diversity Index', y ~ poly(x, 3, raw = TRUE) , 4)

## Save
# bv_ps = grid.arrange(grobs = list(f_n_ps, f_ph_ps, n_ps, ph_ps), ncol = 2, heights = c(2, 1))
# ggsave(filename = paste0(save_path, 'shannon_to_ps_23032023.png'), plot = s,
#        width = 7, height = 3.8, units = "in", limitsize = TRUE)
ggsave(filename = paste0(save_path, 'shannon_to_ps_', date, '.png'), plot = s, height = 6, width = 10, units = "in", dpi = 150)
```

## ROC curves for nugent, ph and amsel

```{r ROC nugent and pH, message = FALSE, warning = FALSE}
## Order
nug_df = read_excel(roc_file, sheet = 'nug', col_names = TRUE)
ph_df = read_excel(roc_file, sheet = 'ph', col_names = TRUE)
bv_df = read_excel(roc_file, sheet = 'bv', col_names = TRUE)
whiff_df = read_excel(roc_file, sheet = 'whiff', col_names = TRUE)
fluid_df = read_excel(roc_file, sheet = 'fluid', col_names = TRUE)
clue_df = read_excel(roc_file, sheet = 'clue', col_names = TRUE)

## Plots
nug_roc = roc_plot(nug_df, nug_df$fpr, nug_df$tpr, nug_df$shuff_num, 'High Nugent Score')
ph_roc = roc_plot(ph_df, ph_df$fpr, ph_df$tpr, 'High pH') +
         theme(axis.title.y = element_blank())
bv_roc = roc_plot(bv_df, bv_df$fpr, bv_df$tpr, 'Positive Amsels Criteria')+
         theme(axis.title.y = element_blank())

## Save
ggarrange(nug_roc, ph_roc, bv_roc, nrow = 1) %>%
ggexport(filename = paste0(save_path, 'ROC_01012023.png'), width = 1500, height = 400)
while (!is.null(dev.list()))  dev.off()
```

## Amines differences

```{r Amines}
met_df = read_excel(paste0(top_path, 'metabolites_with_ps.xlsx'))
met_df$high_ps = as.factor(met_df$high_ps)
get_met_boxplot = function(df, bin_ps_col, col, colname, lab){
  p <- ggplot(data = df, aes(x = factor(bin_ps_col), y = col), color = 'slategray4') +
  geom_boxplot(lwd = 0.65) +
  labs(x = 'Pseudo-time category', y = paste0(colname, ' (', expression('\u03BC'), 'M)')) +
  scale_x_discrete(labels = c('0' = 'Low', '1' = 'High')) +
  geom_signif(comparisons = list(c("0", "1")), test = "wilcox.test", test.args=list(alternative = "less"),
              map_signif_level = FALSE, textsize = 5, tip_length = 0) + #, family = 'Gafata'
  annotate(geom = 'text', label = lab, x = -Inf, y = Inf, hjust = -0.5, vjust = 1.2, family = 'Gafata', size = 8) +
  # stat_summary(fun.data = stat_box_data, geom = "text", fun.y = mean,
  #                 family = 'Gafata', hjust = 0.5, vjust = 0.9, size = 5) +
  guides(fill=guide_legend(ncol=2)) +
  theme_bw(base_family = 'Gafata') +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1))
  return(p)
}
# p <- ggboxplot(met_df, x = 'bin_ps', y = 'Cadaverine',
#           color = "blue", palette = "jco",
#           add = "jitter")
# print(p)

p1 = get_met_boxplot(met_df, met_df$high_ps, met_df$Cadaverine, 'Cadaverine', 'A')
p1 = p1 +  theme(axis.title.x = element_blank())
p2 = get_met_boxplot(met_df, met_df$high_ps, met_df$Putrescine, 'Putrescine', 'B')
p2 = p2 +  theme(axis.title.x = element_blank())
p3 = get_met_boxplot(met_df, met_df$high_ps, met_df$Spermidine, 'Spermidine', 'C')
p3 = p3 +  theme(axis.title.x = element_blank())
p4 = get_met_boxplot(met_df, met_df$high_ps, met_df$Spermine, 'Spermine', 'D')
p5 = get_met_boxplot(met_df, met_df$high_ps, met_df$Tyramine, 'Tyramine', 'E')
p6 = get_met_boxplot(met_df, met_df$high_ps, met_df$total, 'Total', 'F')


# ggarrange(p1, p2, p3, p4, p5, p6, nrow = 2, ncol = 3) %>%
# ggexport(filename = paste0(save_path, 'metbolites_17052023.png'), width = 1300, height = 700, common.legend = TRUE, legend = "bottom")
# while (!is.null(dev.list()))  dev.off()

print(p6)
```


```{r Menst ps variability}
## Order
menst = get_data(menst_file)

## Function


## Plots


## Save
```

## Bacterial changes along arms

```{r Bacterial species}
## Order
melt_df = bacteria_df
melt_df = melt_df[with(melt_df, order(mt_pseudotime, value)),]
melt_df$round_ps = round(melt_df$mt_pseudotime, digits = 3)
melt_df$mt_pseudotime = factor(as.character(melt_df$mt_pseudotime))
melt_df$value = round(melt_df$value, digits = 2)
melt_df$species <- gsub('_', ' ', melt_df$species)
melt_df$species <- gsub('g ', '',
                   gsub('o ', '', melt_df$species))

cols = c("Lactobacillus crispatus" = "#9d0208", #
         "Lactobacillus gasseri" = "#FFD60A", #
         "Lactobacillus iners" = "#e85d04", #
         "Lactobacillus jensenii" = "#E2D6B8", #
         "Lactobacillus Cluster 1" = "#C9184A", #
         "Lactobacillus psittaci" = "#FF758F", #
         "Gardnerella vaginalis" = "#22577a", #
         "BVAB1" = "#7798AB", #
         "Bifidobacterium" = "#10002b", #
         "Staphylococcus" = "#3c096c", #
         "Prevotella" = "#7b2cbf", #
         "Atopobium" = "#e0aaff", #
         "Ureaplasma" = "#d8f3dc", 
         "Anaerococcus" = "#95d5b2",
         "Megasphaera" = "#52b788", #
         "Aerococcus" = "#2d6a4f", #
         "Mageeibacillus" = "#0F4C37", # 
         "Corynebacterium" = "#edc4b3", #
         "Enterococcus" = "#deab90", #
         "Streptococcus" = "#cd9777", #
         "Parvimonas" = "#b07d62", #
         "Finegoldia" = "#774936",
         "Peptoniphilus" = "#2C2C4E",
         "Gemella haemolysans" = "#4a4e69",
         "Dialister" = "#9a8c98",
         "Sneathia" = "#C9ADA7",
         "Saccharimonadales" = "#f2e9e4",
         "Others" = "#878890")

melt_df$species <- factor(melt_df$species,
                      levels = names(cols))#names(cols))

get_barplot = function(df, title){
  g = ggplot(df, aes(fill = species, y = value, x = mt_pseudotime)) + 
    geom_bar(position = "fill", stat = "identity", width = 1) +
    scale_y_continuous(expand = c(0, 0), labels = seq(0, 100, 25)) +
    scale_x_discrete(labels = function(x) {round(as.numeric(x),2)}, breaks = every_nth(n = 5)) +
    scale_fill_manual(values = cols) +
    labs(x = 'Pseudo-time', y = 'Relative Abundance', title = paste0('', title)) +
    theme_bw(base_family = 'Gafata') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
          text = element_text(size = 20))

  return(g)
}

## Plot
g0 = get_barplot(melt_df, 'legend') +
      theme(legend.position = 'bottom',
            legend.title = element_blank(), legend.text=element_text(size = 35)) +
      guides(fill = guide_legend(override.aes = list(size = 0.15), nrow = 7))

## Save
ggsave(paste0(save_path, 'LEGEND_bacterial_abundance_01012023.png'), plot = g0,
       width = 12, height = 8, units = 'in')
while (!is.null(dev.list()))  dev.off()

## Plots
g1 = get_barplot(subset(melt_df, branch == 'I'), 'I') +
      theme(legend.position = 'none', axis.title.x = element_blank())
# 
g2 = get_barplot(subset(melt_df, branch == 'II'), 'II') +
      theme(legend.position = 'none', axis.title.x = element_blank(), axis.title.y = element_blank())

g3 = get_barplot(subset(melt_df, branch == 'III'), 'III') +
      theme(legend.position = 'none', axis.title.x = element_blank(), axis.title.y = element_blank())

g4 = get_barplot(subset(melt_df, branch == 'V'), 'V') +
      theme(legend.position = 'none')

g5 = get_barplot(subset(melt_df, branch == 'IV-A'), 'IV-A') +
      theme(legend.position = 'none', axis.title.y = element_blank())

g6 = get_barplot(subset(melt_df, branch == 'IV-B'), 'IV-B') +
      theme(legend.position = 'none', axis.title.y = element_blank())

## Save
ggarrange(g1, g2, g3, g4, g5, g6, nrow = 2, ncol = 3) %>%
ggexport(filename = paste0(save_path, 'bacterial_abundance_0.2_01012023.png'), width = 1200, height = 700,             common.legend = TRUE, legend = "bottom")
while (!is.null(dev.list()))  dev.off()
```
## BC on X to euclidean on PCA

```{r}
bc_meta = get_data(paste0(top_path, 'ps_res_BC_X_25062023.xlsx'), sheet_meta, coltypes_vec)
bc_umap = get_data(paste0(top_path, 'ps_res_BC_X_25062023.xlsx'), sheet_umap, c("text", "numeric", "numeric"))
meta = get_data(paste0(top_path, 'ps_res_16052023.xlsx'), sheet_meta, coltypes_vec)
umap = get_data(paste0(top_path, 'ps_res_16052023.xlsx'), sheet_umap, c("text", "numeric", "numeric"))
bc_all_meta = get_meta_order(bc_meta)
bc_all_meta = bc_all_meta %>% 
  left_join(bc_umap %>% rownames_to_column("run"), by = c("run"))
all_meta = get_meta_order(meta)
all_meta = all_meta %>% 
  left_join(bc_umap %>% rownames_to_column("run"), by = c("run")) 

scatter_df = bc_all_meta %>% rename(mt_pseudotime_bc = mt_pseudotime) %>% select(c('sampleID', 'mt_pseudotime_bc')) %>%
             left_join(all_meta)

get_scatter_comparison = function(df, x, y, col, xlab, ylab){
  s = ggplot(data = df) +
      geom_point(aes(x = x, y = y, fill = col), shape = 21, alpha = 0.5, size = 1.5) +
      labs(x = xlab, y = ylab) +
      theme_bw(base_family = 'Gafata') +
      geom_abline() +
      scale_fill_manual(values = community_colors, name = 'subCST') +
      theme(text = element_text(size = 30))
  return(s)
}

s = get_scatter_comparison(scatter_df, scatter_df$mt_pseudotime_bc, scatter_df$mt_pseudotime, scatter_df$subCST, 'Pseudotime BC based', 'Pseudotime euclidean based')

ggsave(paste0(save_path, 'euclidean_to_bc_25062023.png'), plot = s,
       width = 5, height = 3, units = 'in')
```


